Neural models for computer-assisted musical orchestration: a preliminary study

### 2020-05-08 paper/latex/paper.tex
fix distance eqn
Add baseline table

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.

@@

For a slight modification on this setting, we no longer attempted to identify which instrument was present if it was not one of the two instruments whose pitch class was being identified. Instead, the classifier would simply identify that an instrument that was not one of the two was present. Since this is a simpler problem, it lead to increased accuracies. Flute and Violin went from 30.2\% to 38.8\%, Oboe and French Horn from 27.3\% to 39.9\%, and Trombone and Bass Tuba from 2.2\% to 3.5\%. \textit{(I think we should maybe put a table of this data instead of listing it all out)} Random Forest performed significantly worse for this setting. Flute and Violin had an accuracy of 9.8\% and Oboe and French Horn was 17.5\%. For this reason, we stopped testing with Random Forest and continued with SVM only.

@@

@@

For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets. The distance metric we used was

$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$

$$\delta_{k1} = 
\begin{cases}
1, \text{if   } x_k > \tilde{x}_k \\
0, \text{otherwise}
\end{cases} $$

$$ \delta_{k2} = 1 - \delta_{k1}$$

>>>>

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors. We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For Random Forest, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments. All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40. Each sample is a combination of one or more instruments and is four seconds in length. The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample. We chose MFCCs because of their ability to describe the timbre of a sample using a relatively small number of coefficients.

@@

For a slight modification on this setting, we no longer attempted to identify which instrument was present if it was not one of the two instruments whose pitch class was being identified. Instead, the classifier would simply identify that an instrument that was not one of the two was present. Since this is a simpler problem, it lead to increased accuracies. The accuracies from this experiment are outlined in Table 1.

@@

\begin{table}[h!]
  \begin{center}
    \label{tab:table1}
    \begin{tabular}{c|c|c|c} 
    	  \hline
      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Accuracy} & \textbf{RF Accuracy}\\
      \hline
      Violin & Flute & 38.8\% & 9.8\% \\
      \hline
      Violin & Trumpet & 33.8\% & 9.1\% \\
      \hline
      Violin & Cello & 34.8\% & 6.3\% \\
      \hline
      Cello & Viola & 32.1\% & 5.8\% \\
      \hline
      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Comparison of results between SVM and Random Forest. Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment. For the samples drawn from one of the two instruments, the pitch class of that sample is identified. For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present. In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.}
\end{table}

@@

For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets. The distance metric we use is defined in \eqref{distance}.

\begin{equation}\label{distance}
d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
\end{equation}

$$
\delta_{k1} = 
\begin{cases}
1, \text{if   } x_k > \tilde{x}_k \\
0, \text{otherwise}
\end{cases} 
$$

$$
\delta_{k2} = 
\begin{cases}
1, \text{if   } x_k < \tilde{x}_k \\
0, \text{otherwise}
\end{cases} 
$$
